---
title: "readme"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prerequisites

Before installing FLfse, the following non-CRAN R packages should be installed:

* stockassessment (available from https://github.com/fishfollower/SAM)
* spict (available from https://github.com/mawp/spict/)
* FLR packages FLCore and ggplotFL (see http://www.flr-project.org/#install)

## Install FLfse

FLfse can be installed with
```{R, eval = FALSE}
devtools::install_github("shfischer/FLfse/FLfse")
```
## Showcase functionality
load package
```{r}
library(FLfse)
```
The package includes example data for some stocks, see
```{R}
data(package = "FLfse")
```
Currently, this includes North Sea cod and Irish Sea plaice.

# Run stock assessments
## SAM
Fit SAM to North Sea cod data, with default parametrizations:
```{R}
fit <- FLR_SAM(stk = cod4_stk, idx = cod4_idx)
```
use the SAM assessment configuration as used by ICES WGNSSK 2018:
```{r}
### fit SAM model
fit <- FLR_SAM(stk = cod4_stk, idx = cod4_idx, conf = cod4_conf_sam)
### check convergence
fit$opt$convergence

### The result of FLR_SAM() is an object of class "sam". 
### All methods defined in the stockassessment package can be used on it.
### summary table
summary(fit)
### plot model
plot(fit)
### plot catch
catchplot(fit)
```
The resulting "sam" object can then be converted into an FLStock object:
```{R}
cod4 <- SAM2FLStock(fit, cod4_stk)
### plot with ggplotFL
library(ggplotFL)
plot(cod4)
```

Irish Sea plaice is also included, as run by ICES WGCSE 2017:
```{R, eval = FALSE}
ple_iris <- FLR_SAM(stk = ple7a_stk, idx = ple7a_idx, conf = ple7a_conf_sam)
### check convergence
ple_iris$opt$convergence
### results
summary(ple_iris)
plot(ple_iris)
```

## SPiCT
Another assessment method implemented is the Surplus Production in Continuous Time (SPiCT) model (Pedersen & Berg, 2017).

```{R}
library(spict)
### fit SPiCT to Irish Sea plaice
fit <- FLR_SPiCT(stk = ple7a_stk, idx = ple7a_idx)
### check results
fit
plot(fit)

### pass additional configuration, set time step to 1 per year
conf <- list(dteuler = 1)
fit2 <- FLR_SPiCT(stk = ple7a_stk, idx = ple7a_idx, conf = conf)
fit2
```

