---
title: "FLfse"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

An R package for providing an interface between the SAM stock assessment (SAM, https://github.com/fishfollower/SAM) and the Fisheries Library in R (FLR, https://flr-project.org/).

The package provides methods for running SAM with FLR objects (`FLR_SAM`) and functions useful for conducting a Management Strategy Evaluation (MSE). This includes methods for turning the output of a SAM model fit into FLR's `FLStock` (`SAM2FLStock`) and for creating simulation replicates based on SAM's variance-covariance matrix (`SAM_uncertainty`).

## Prerequisites

Before installing FLfse, the following non-CRAN R packages should be installed:

* stockassessment (available from https://github.com/fishfollower/SAM)
* FLR packages FLCore and ggplotFL (see http://www.flr-project.org/#install)

## Install FLfse

FLfse can be installed with
```{R, eval = FALSE}
remotes::install_github("shfischer/FLfse/FLfse")
```
## Showcase functionality
Load package
```{r}
library(FLfse)
```
The package includes example data for some stocks, see
```{R, eval = FALSE}
data(package = "FLfse")
```
Among others, this includes the 2018 assessments for North Sea cod, North Sea haddock and North Sea whiting and the 2019 assessment for North Sea herring.

### Run SAM stock assessments
#### North Sea cod
Fit SAM to North Sea cod data from ICES WGNSSK 2018, with default parameterisations:
```{R, message = FALSE, warning = FALSE}
fit <- FLR_SAM(stk = cod4_stk, idx = cod4_idx)
```
use the SAM assessment configuration as used by ICES WGNSSK 2018:
```{r}
### fit SAM model
fit <- FLR_SAM(stk = cod4_stk, idx = cod4_idx, conf = cod4_conf_sam)
### check convergence
fit$opt$convergence

### The result of FLR_SAM() is an object of class "sam". 
### All methods defined in the stockassessment package can be used on it.
### summary table
tail(summary(fit))
### plot model
plot(fit)
```
The resulting "sam" object can then be converted into an FLStock object:
```{R, message = FALSE, warning = FALSE}
cod4 <- SAM2FLStock(fit, cod4_stk)
### plot with ggplotFL
library(ggplotFL)
plot(cod4)
```

#### North Sea herring

The 2019 ICES HAWG assessment for North Sea herring can be replicated with `FLfse`.
To get this assessment running, a different branch ("components") of the `stockassessment` package needs to be installed.
The following code snippet shows how to do this without interfering with the global packages (please restart R if another version of stockassessment is currently loaded):

```{R, eval = FALSE}
### 1st: add a temporary R package library 
.libPaths(c(tempdir(), .libPaths()))
### 2nd: install stockassessment branch "components"
devtools::install_github("fishfollower/SAM/stockassessment", ref = "components")

### now, the assessment can be run:
library(FLfse)
library(stockassessment)
### replicate HAWG assessment
fit <- FLR_SAM(stk = her4_stk, idx = her4_idx, conf = her4_conf_sam, 
                    conf_full = TRUE, NA_rm = FALSE)
### check results
fit
plot(fit)
summary(fit)
### convert into FLStock
stk <- SAM2FLStock(fit, her4_stk)
```

### Create simulation replicates from SAM model fit

We can use SAM's variance-covariance matrix to create simulation replicates (e.g. for MSE).

```{r, message = FALSE, warning = FALSE}
### fit SAM to North Sea cod
fit <- FLR_SAM(stk = cod4_stk, idx = cod4_idx, conf = cod4_conf_sam)
### create simulation replicates
set.seed(1)
unc <- SAM_uncertainty(fit = fit, n = 10)
### turn into FLStock
stk <- SAM2FLStock(fit, cod4_stk)
### insert replicates for stock numbers, catch numbers and fishing mortality
stk <- propagate(stk, 10)
stock.n(stk) <- unc$stock.n ### insert stock numbers
stock <- computeStock(stk) ### update stock biomass
catch.n(stk)[, ac(1963:2017)] <- unc$catch.n ### insert catch numbers
catch(stk) <- computeCatch(stk) ### update total catch
harvest(stk) <- unc$harvest ### insert fishing mortality at age
units(harvest(stk)) <- "f"
### plot stock and replicates
plot(stk, iter = 1:10)
```
